---
title: "Review"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

1.  **Clone the Homework Repository**\
    Clone the GitHub repository containing the homework file to your **local machine**.

2.  **Rename the File**\
    Rename the `.qmd` file from `hw01.qmd` to: `hw01_yourusername.qmd`.\

For example: `hw01_semiyari.qmd`

### Working on Your Homework

-   As you complete each part of the homework:

    -   **Stage** and **commit** your `.qmd` file *after each question/sub-question*, such as `1.1`, `1.2`, etc.\
    -   **Render** your `.qmd` file to **HTML only** (no PDF or Word formats).\
    -   **Push** your updates to GitHub after completing each **main question**, such as `1.0`, `2.0`, etc.

> Committing and pushing to GitHub **are graded components** of this assignment.

## Final Checks Before Submission

-   **Re-render** to HTML before your final commit and push.\

-   **Check your output:**\

    -   Make sure **all images display correctly**.\
    -   Avoid including **excessive output or large datasets** unless asked.

## Final Submission

Once you have finished and pushed your final version:

-   Copy the link of your **GitHub Pages site** (where the rendered HTML is published).
-   **Post the link** to the Canvas assignment submission page.

```{r}
# Install pacman if it's not already installed
if (!require("pacman")) install.packages("pacman")

# Use pacman to install (if needed) and load the packages
pacman::p_load(pacman, tidyverse,
               nycflights13)
```

-   Consider the data in the nycflights13 package.

```{r}
library(nycflights13)

head(flights)
```


-   **Problem 2**: Is there a relationship between the age of a plane and its delays?

```{r}
library(tidyverse)
library(nycflights13)

flights_with_age <- flights %>%
  left_join(planes, by = "tailnum") %>%
  mutate(plane_age = 2013 - year.y)  

flights_clean <- flights_with_age %>%
  filter(!is.na(dep_delay), !is.na(plane_age))

flights_clean %>%
  ggplot(aes(x = plane_age, y = dep_delay)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Relationship Between Plane Age and Departure Delay",
    x = "Plane Age (years)",
    y = "Departure Delay (minutes)"
  )

```
```{r}
model <- lm(dep_delay ~ plane_age, data = flights_clean)
summary(model)

```
The linear regression shows a statistically significant but very weak relationship between plane age and departure delay. The coefficient for plane_age is -0.097, meaning that for each additional year of a plane’s age, the expected departure delay decreases by about 0.1 minutes. While this result is statistically significant (p < 0.001), the R-squared value is just 0.0002, indicating that plane age explains almost none of the variation in delay times. In short, the relationship is real but not meaningful in practice.


-   **Problem 3**: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.
