---
title: "Functions-Iterations-Dates-purr"
format: html
editor: visual
---

## Instructions

Complete the following exercises using the `lubridate` and `purrr` packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

## Exercise Functions:

### Question 1

Write a function (using r code and structure demonstrated in class) to calculate a z score for a given observed value, a mean, and a standard deviation value. And then use your function to find a z score for the following problem. (Research the internet to find the formula used to calculate a z score)\
Observed value = 25.77, mean = 23.54, standard deviation = 2.442

```{r}
z_score <- function(observed, mean, sd) {
  (observed - mean) / sd
}

z_score(25.77, 23.54, 2.442)

```

### Question 2

Write a function to calculate the natural log of a number multiplied by the common log of the same number divided by the cube root of a given prime number, $\dfrac{\ln x \times \log x }{\sqrt[3]{p}}$.\
Use your function to find the answer if the number to be used for both log expressions is 32 and the given prime number is 11, $x = 32, \;\;p=11$.\
Also use R code to round your answer to the nearest tenth

```{r}
log_math <- function(x, p) {
  result <- (log(x) * log10(x)) / (p^(1/3))
  round(result, 1)  
}

log_math(32, 11)

```

### Question 3

Use and show R coding to calculate the standard deviation for each variable of the data table `mtcars` using the `Special For Loop Method`, which includes three components:

-   Output – an object to store the results

-   Sequence – the set of items to loop through

-   Body – the operations performed in each iteration

```{r}
std_devs <- c()

for (var in names(mtcars)) {
  std_devs[var] <- sd(mtcars[[var]])
}

std_devs

```

### Question 4:

Write only one line of the code to return only the leap years fo years 2024, 3021, 2000 or 1800, 800, 5050, 600 and 7156. using subsetting with `[]` and without using built-in functions.

```{r}
c(2024, 3021, 2000, 1800, 800, 5050, 600, 7156)[(c(2024, 3021, 2000, 1800, 800, 5050, 600, 7156) %% 4 == 0 & 
    c(2024, 3021, 2000, 1800, 800, 5050, 600, 7156) %% 100 != 0) | 
                                                  c(2024, 3021, 2000, 1800, 800, 5050, 600, 7156) %% 400 == 0]

```

## Advanced Date Manipulation with lubridate

### Question 5:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

```{r}
library(lubridate)

date_seq <- seq.Date(from = as.Date("2015-01-01"),
                     to = as.Date("2025-12-31"),
                     by = "2 months")

data.frame(
  date = date_seq,
  year = year(date_seq),
  quarter = quarter(date_seq),
  iso_week = isoweek(date_seq)
)

```

### Question 6:

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.


```{r}
sample_dates <- as.Date(c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05"))

month_diff <- interval(sample_dates[-length(sample_dates)], sample_dates[-1]) %/% months(1)
week_diff <- difftime(sample_dates[-1], sample_dates[-length(sample_dates)], units = "weeks")

data.frame(
  start_date = sample_dates[-length(sample_dates)],
  end_date = sample_dates[-1],
  months_apart = month_diff,
  weeks_apart = as.numeric(week_diff)
)
 #commit done
```

### Question 7: Higher-Order Functions with purrr
Using `map()` and `map_dbl()`, compute the **mean**, **median**, and **standard deviation** for each numeric vector in the following list:
  
  ``` r
num_lists <- list(c(4, 16, 25, 36, 49), c(2.3, 5.7, 8.1, 11.4), c(10, 20, 30, 40, 50))
```

```{r}
library(purrr)

num_lists <- list(c(4, 16, 25, 36, 49), c(2.3, 5.7, 8.1, 11.4), c(10, 20, 30, 40, 50))

means <- map_dbl(num_lists, mean)
medians <- map_dbl(num_lists, median)
sds <- map_dbl(num_lists, sd)

data.frame(
  mean = means,
  median = medians,
  sd = sds
)

```


### Question 8: Combining lubridate and purrr
Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` r
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")
```

```{r}
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")

safe_parse <- possibly(function(x) parse_date_time(x, orders = c("ymd", "Ymd", "d-b-Y", "Y/m/d")), NA)

map(date_strings, ~ month(safe_parse(.x), label = TRUE, abbr = FALSE))

```


## Submission

-   Save your Quarto file after completing each problem, render it, then stage and commit the changes.\

-   Include well-commented code and clear explanations.\

-   Make at least **one commit per question**.\

-   Submit the rendered file to **Canvas** and push it to your **GitHub repository**.\

-   Provide the link to your GitHub repository in the **Canvas gradebook** as a text submission.\

🔹 **Each problem is worth 10 points.**\

🔹 You will receive an additional **20 points** if you correctly push your completed file to GitHub and make at least one commit for each question.\

**Good luck!**
