---
title: "Homework Assignment: Sentiment Analysis of Emma"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

**Title**: Exploring Joyful Language in Jane Austen’s *Emma* using Tidytext

### **Instructions**

Complete the following exercises using the appropriate packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

#### **Objective**:

Use the **`tidytext`** package and three different **sentiment lexicons** (`nrc`, `afinn`, `bing`) to explore **positive/joyful words** in *Emma* by Jane Austen. You will tokenize the text, apply sentiment filters, visualize frequent sentiment words using `ggplot2`, and create a word cloud.

### **Tasks**:

```{r}
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(stringr)
```


1.  **Data Preparation**

    -   Load the `austen_books()` dataset from the **`janeaustenr`** package.
    -   Group by book and detect chapter boundaries using regex.
    -   Create `linenumber` and `chapter` columns.

```{r}
emma_data <- austen_books() %>%
  filter(book == "Emma") %>%                            # keep only Emma
  group_by(book) %>%
  mutate(
    linenumber = row_number(),                          # line number within book
    chapter = cumsum(str_detect(                        # detect chapter start
      text,
      regex("^chapter [\\divxlc]+$", ignore_case = TRUE)
    ))
  ) %>%
  ungroup()

head(emma_data, 10)

```

2.  **Tokenization**

    -   Use `unnest_tokens()` to tokenize text into individual words.
```{r}
emma_tokens <- emma_data %>%
  unnest_tokens(word, text) %>%          # break text into single words
  filter(!str_detect(word, "^[0-9]+$")) %>%  # remove numbers
  anti_join(stop_words, by = "word")     # remove common stop words

# View the first few tokenized words
head(emma_tokens, 10)

```


3.  **Sentiment Analysis**

    -   Filter joy/positive words from **each** of the three sentiment lexicons:

        -   `nrc` (joy)
        -   `afinn` (positive scores ≥ 1)
        -   `bing` (positive)
```{r}
library(tidytext)
library(dplyr)

nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy") %>%
  distinct(word)            # keep unique joy words

afinn_pos <- get_sentiments("afinn") %>%
  filter(value >= 1) %>%    # positive scores only
  select(word, value)       # keep value for optional weighting later

bing_pos <- get_sentiments("bing") %>%
  filter(sentiment == "positive") %>%
  distinct(word)

```

    -   Join each with *Emma*'s text and:
    
```{r}
emma_nrc_counts <- emma_tokens %>%
  inner_join(nrc_joy, by = "word") %>%
  count(word, sort = TRUE)

# AFINN positive (>=1)
emma_afinn_counts <- emma_tokens %>%
  inner_join(afinn_pos %>% select(word), by = "word") %>%  # keep just word to count frequency
  count(word, sort = TRUE)

# Bing positive
emma_bing_counts <- emma_tokens %>%
  inner_join(bing_pos, by = "word") %>%
  count(word, sort = TRUE)

# quick peek
list(
  nrc_head   = head(emma_nrc_counts, 10),
  afinn_head = head(emma_afinn_counts, 10),
  bing_head  = head(emma_bing_counts, 10)
)

```


        -   Count word frequency.
        -   Filter for frequently occurring words (`n > 50`).
        -   Visualize using a **bar chart** (`ggplot2`) and a **word cloud** (`wordcloud`).
        
```{r}
thr <- 50

nrc_top   <- emma_nrc_counts   %>% dplyr::filter(n > thr)
afinn_top <- emma_afinn_counts %>% dplyr::filter(n > thr)
bing_top  <- emma_bing_counts  %>% dplyr::filter(n > thr)

plot_bar <- function(df, title) {
  ggplot(df, aes(x = reorder(word, n), y = n)) +
    geom_col() +
    coord_flip() +
    labs(x = NULL, y = "Count", title = title)
}

print(plot_bar(nrc_top,   "Emma — NRC Joy (n > 50)"))
print(plot_bar(afinn_top, "Emma — AFINN Positive (n > 50)"))
print(plot_bar(bing_top,  "Emma — Bing Positive (n > 50)"))

```
```{r}
library(wordcloud)
set.seed(123)  # reproducible layout

wordcloud(nrc_top$word,   nrc_top$n,   max.words = 100, scale = c(1, 0.3))
wordcloud(afinn_top$word, afinn_top$n, max.words = 100, scale = c(1, 0.3))
wordcloud(bing_top$word,  bing_top$n,  max.words = 100, scale = c(1, 0.3))


```
        

4.  **Push to GitHub**

    -   Push your complete R script (`.R` file) to a **GitHub repository**.

    -   Your script should include:

        -   Data wrangling
        -   Sentiment filtering and joins
        -   Visualization code

    -   You must include **at least 5 meaningful commits** to document your workflow.

------------------------------------------------------------------------

## **Rubric 100 Points**

| Category                                       | Points  |
|------------------------------------------------|---------|
| Step 1: Data wrangling (linenumber, chapter)   | 10      |
| Step 2: Tokenization                           | 10      |
| Step 3a: NRC sentiment analysis + plots        | 10      |
| Step 3b: AFINN sentiment analysis + plots      | 10      |
| Step 3c: BING sentiment analysis + plots       | 10      |
| Code readability, structure, and comments      | 10      |
| Word cloud for each lexicon                    | 20      |
| GitHub submission                              |         |
| Commit history (at least 5 meaningful commits) | 20      |
| **Total**                                      | **100** |

Good Luck!
